Laplacian envelope

### 2021-10-01 Notes/laplace.tex
written about area maxers KKT

This proof follows that of \cite{Lawler1988cheeger}.
We can define two inner-products that will be of use.
The first is the usual inner product of two vectors (or functions on state space in the continuum case) using the steady state distribution as the measure:
%
\begin{equation}\label{eq:innereq}
  \av{\mathbf{u}, \mathbf{v}} = \sum_i \eqm_i u_i^\ast v_i ,
  \qquad
  \nrm{\mathbf{v}} = \sqrt{\av{\mathbf{v}, \mathbf{v}}} .
\end{equation}
%
The second is an inner product of rank 2 tensors (functions of pairs of points on state space) using the steady-state flux as the measure:
%
\begin{equation}\label{eq:innerflux}
  \aav{\mathbf{F}, \mathbf{G}} = \sum_{i \neq j} \Fm_{ij} F_{ij}^\ast G_{ij} ,
  \qquad
  \nnrm{\mathbf{F}} = \sqrt{\aav{\mathbf{F}, \mathbf{F}}}.
\end{equation}
%
The omission of the diagonal terms leaves only non-negative terms.

@@

The other quantity of interest in \cref{eq:snrcoeffs} is the amplitude of the eigenmodes.
In analogy, we define the following functional (for real \(\mathbf{u}\))
%
\begin{equation*}
  \CI(\mathbf{u}) = \frac{(\eq \enc \mathbf{u}) (\mathbf{u}^\dag \w)}{r\mathbf{u}^\dag \mathbf{u}}
      = \frac{\sum_{i \neq j} \eqm_i \encm_{ij} (u_j - u_i) \sum_k \eqm_k u_k \wm_k} 
             {r\sum_{l} \eqm_l u_l^2} .
\end{equation*}
%
Following \cite{Lawler1988cheeger}, we can write
%
\begin{equation*}
  u_j - u_i = \intd[_{u_i}^{u_j}]{\alpha} 
      = \intd[_{-\infty}^{\infty}]{\alpha} 
        \brk{\vartheta(u_j - \alpha) \vartheta(\alpha - u_i) 
           - \vartheta(u_i - \alpha) \vartheta(\alpha - u_j)} ,
\end{equation*}
%
where \(\vartheta\) is the Heaviside step-function.
Then 
%
\begin{equation*}
  \begin{aligned}[t]
  \eq \enc \mathbf{u} &= \intd[_{-\infty}^{\infty}]{\alpha}  \brk{
        \Fm\pot_{\CA_\alpha \CA_\alpha^c} - \Fm\dep_{\CA_\alpha \CA_\alpha^c} 
      - \Fm\pot_{\CA_\alpha^c \CA_\alpha} + \Fm\dep_{\CA_\alpha^c \CA_\alpha} }
  %
  \; \text{where} 
    \\
    &= \intd[_{-\infty}^{\infty}]{\alpha}  \brk{
        \Fm_{\CA_\alpha \CA_\alpha^c} + \Fm_{\CA_\alpha^c \CA_\alpha} 
        - 2\Fm\dep_{\CA_\alpha \CA_\alpha^c} - 2\Fm\pot_{\CA_\alpha^c \CA_\alpha} } ,
  \end{aligned}
  %
  \begin{aligned}[t]
    \CA_\alpha &= \set{i}{u_i < \alpha} , 
    \\
    \Fm^\nu_{\CA \CB} &= \sum_{\substack{i \in \CA \\ j \in \CB}} 
                        \eqm_i \Wm^\nu_{ij} .
  \end{aligned}
\end{equation*}
%
Note that the first two terms of the integrand are of the form \( \Fm_{\CA \CB} = \Fm_{\CA \CB}\pot + \Fm_{\CA \CB}\dep \).
In exactly the same way as the proof of \cref{eq:initbnd}, we can argue that the two terms with the negative signs are indeed non-positive as they do not involve the negative diagonal elements of~\( \Fm_{ij}^\nu \).
Therefore the other terms provide an upper bound:
%
\begin{equation*}
\begin{aligned}
  \CI(\mathbf{u}) 
      &\leq \frac{\av{\mathbf{u},\w}}{r\nrm{\mathbf{u}}^2} 
        \intd[_{-\infty}^{\infty}]{\alpha}  \brk{
        \Fm_{\CA_\alpha \CA_\alpha^c} + \Fm_{\CA_\alpha^c \CA_\alpha} } 
  \\  &= \frac{\av{\mathbf{u},\w}}{r\nrm{\mathbf{u}}^2} 
        \intd[_{-\infty}^{\infty}]{\alpha} \sum_{i \neq j} \Fm_{ij} \brk{
            \vartheta(u_j - \alpha) \vartheta(\alpha - u_i) 
          + \vartheta(u_i - \alpha) \vartheta(\alpha - u_j)
        }
  \\  &= \frac{\av{\mathbf{u},\w}}{r\nrm{\mathbf{u}}^2} 
      \sum_{i \neq j} \Fm_{ij} \abs{u_j - u_i}
  \\  &= \frac{\av{\mathbf{u},\w} \aav{\abs{\boldsymbol{\Delta}\mathbf{u}}, \onem}}
              {r\nrm{\mathbf{u}}^2}.
\end{aligned}
\end{equation*}
%
Now we can apply the Cauchy-Schwartz inequality to both factors to find
%
\begin{equation*}
  \CI(\mathbf{u}) \leq \frac{\nrm{\w} \nnrm{\boldsymbol{\Delta}\mathbf{u}} \nnrm{\onem}}
                            {r\nrm{\mathbf{u}}} .
\end{equation*}
%
We also have \( \nrm{\w}^2 = \sum_i \eqm_i \wm_i^2 = 1 \).
Furthermore, using \cref{eq:constri}, \( \nnrm{\onem}^2 = \sum_{i \neq j} \eqm_i \MMm_{ij} \leq r \sum_i \eqm_i = r\).
Then, remembering \cref{eq:numrange},
%
\begin{equation}\label{eq:sqrtnumrange}
  \CI(\mathbf{u}) \leq \sqrt{\frac{2\av{\mathbf{u}, -\MM \mathbf{u}}}{r\av{\mathbf{u}, \mathbf{u}}}}
      = \sqrt{2\mathcal{W}(-\frg;\mathbf{u})/r}.
\end{equation}
%
Now, simply choosing \(\mathbf{u} = \evr_a \), one of the orthonormal eigenvectors, we find a bound on the eigenmodes:
%
\begin{equation}\label{eq:sqrtbound}
  \text{assuming} \quad 
  \eqm_i \frgm_{ij} = \eqm_j \frgm_{ji},
  \qquad \text{we have} \quad
  \CI_a \sqrt{r\tau_a} \leq \sqrt{2}.
\end{equation}
%
Numerical maximisation of the quantity above produces the same result as in \cref{sec:initial}:
the two-state model with deterministic transitions.
This model has \(\CI_1 = \tau_1 = 1\), suggesting that there's a factor of \(\sqrt{2}\) looseness in this bound.

@@

Suppose there are two states, \(i,j\), with nonzero transitions from \(m\) for one \(\nu\).
Both must statisfy \cref{eq:oneexit} for the same \(\kktm^\nu_{mm}\), so
%
\begin{equation}\label{eq:twoexits}
  \Wm^\nu_{mi}, \Wm^\nu_{mj} > 0
  \qquad \implies \qquad
  \dgnm^\mu_{m,i} = \dgnm^\nu_{m,j}
  \quad \text{or}\quad
  \dgnm_{m,i} - \dgnm_{m,j} = - (-1)^\nu (\etwm_i - \etwm_j).
\end{equation}
%
Furthermore, for another state \(k\) with zero transition rate from \(m\) for the same \(\nu\):
%
\begin{equation}\label{eq:nonexit}
\begin{aligned}
  \Wm^\nu_{mi} > 0, \Wm^\nu_{mk} = 0 
  \;\;\; \implies \;\;\;
  \kktm^\nu_{mk} = \eqm_m (\dgnm^\nu_{m,k} - \dgnm^\nu_{m,i})
  \;\; \implies \;\;
  \dgnm^\nu_{m,k} &\geq \dgnm^\nu_{m,i} \\
  \text{or}\quad
  \dgnm_{m,k} - \dgnm_{m,i} &\geq - (-1)^\nu (\etwm_k - \etwm_i).
\end{aligned}
\end{equation}
%
If both \(\nu\) have nonzero rates from \(m\) to \(i\) and \(j\):
%
\begin{equation}\label{eq:twodoubleexits}
  \Wm\pot_{mi}, \Wm\dep_{mi}, \Wm\pot_{mj}, \Wm\dep_{mj} > 0
  \qquad \implies \qquad
  \dgnm_{m,i} = \dgnm_{m,j} 
  \quad \text{and} \quad
  \etwm_i = \etwm_j.
\end{equation}
%
The consequences of these inequalities are depicted in \cref{fig:planes}(\ref{fig:triangle_spacing}).
We see that \(n\) must be a global minimum of \(\dgn_m\), and if there are multiple transitions of a single type (potentiating or depressing) the targets are constrained to lie on a single line.
Even if there is no state \(n\) that is the target of both potentiating and depressing transitions, there will be two lines at \(45^\circ \) passing through any targets of potentiating/depressing transitions that demarcate the forbidden zone.

@@

\begin{figure}[ht]
\begin{center}
\begin{myenuma}
  \item \label{fig:triangle_spacing} \aligntop{ \hspace{-2em}%
  \includegraphics[width=0.4\linewidth]{spacing_triangle.svg}}
  \hspace{0.2em}
  \item \label{fig:scaling_plane}\aligntop{ \hspace{-1em}%
  \includegraphics[width=0.4\linewidth]{scaling_plane.svg}}
\end{myenuma}
\caption[Depicting constraints on topology]{Depicting constraints on topology.
(\ref*{fig:triangle_spacing}) The consequences of \cref{eq:oneexit,eq:doubleexit,eq:twoexits,eq:twodoubleexits,eq:nonexit} in the \( (\dgnm_{m,l},-\etwm_l) \)-plane, 
  with \(i\) a state with a nonzero depressing transition from \(m\), 
  \(j\) a state with a nonzero potentiating transition, 
  \(n\) a state with both depressing and potentiating transitions 
  and \(k\) a state with no nonzero transitions from \(m\).
  The positions of \(m\) and \(k\) are not yet constrained, other than being in the upper quadrant \wrt \(n\).
(\ref*{fig:scaling_plane}) Scaling of optimal models in the \((s,r)\)-plane.
  Optimal models on grey ``rays'' are related by scaling, and therefore have the same topology.
  The line at \(r = r_0\) is where we need the solutions.
  The line at \(s = \epsilon\) is where we can determine the topology.
\label{fig:planes}}
\end{center}
\end{figure}

@@

We have the following invariances:
%
\begin{equation*}
  A(\lambda s, \lambda \W) = A(s, \W),
  \qquad
  \CL(\lambda s, \lambda r, \lambda \W, \lambda\inv \kkt) = \CL(s, r, \W, \kkt).
\end{equation*}
%
or in differential form:
%
\begin{equation*}
  s \pdiffc<\Wm>{A}{s} + \sum_{\nu i j} \Wm^\nu_{ij} \pdiffc<s>{A}{\Wm^\nu_{ij}} 
  = 
  s \pdiffc<r \Wm \mathrlap{\kktm}>{\CL}{s} 
    + r \pdiffc<s \Wm \mathrlap{\kktm}>{\CL}{r} 
    + \sum_{\nu i j} \brk{\Wm^\nu_{ij} \pdiffc<s r \mathrlap{\kktm}>{\CL}{\Wm^\nu_{ij}} 
                      - \kktm^\nu_{ij} \pdiffc<s r {\Wm}>{\CL}{\kktm^\nu_{ij}}} = 0.
\end{equation*}
%
In \cref{eq:kktsol} we have an expression for the KKT multipliers at the maximum in terms of \(s\) and \(\W \).
We denote this expression by \( \kkt^\ddagger(s, \W) \).
If we denote the solution to the maximisation problem by \(\W^\ast(s,r), \kkt^\ast(s,r)\), we have \( \kkt^\ast(s,r) = \kkt^\ddagger(s, \W^\ast(s,r)) \) and
%
\begin{equation}\label{eq:optmodelscale}
\begin{aligned}
  \kkt^\ddagger(\lambda s, \lambda \W) 
    &= \lambda\inv \kkt^\ddagger(s, \W), &
  \qquad \text{or} \qquad
  s \pdiffc<\Wm>{\kkt^\ddagger}{s} 
      + \sum_{\nu i j} \Wm^\nu_{ij} \pdiffc<s>{\kkt^\ddagger}{\Wm^\nu_{ij}} 
    &= -\kkt^\ddagger, \\
  \W^\ast(\lambda s, \lambda r) 
    &= \lambda \W^\ast(s,r), &
  s \pdiffc<r>{\W^\ast}{s} + r \pdiffc<s>{\W^\ast}{r} 
    &= \W^\ast, \\
  \kkt^\ast(\lambda s, \lambda r) &= \lambda\inv \kkt^\ast(s,r), &
  s \pdiffc<r>{\kkt^\ast}{s} + r \pdiffc<s>{\kkt^\ast}{r} 
    &= -\kkt^\ast.
\end{aligned}
\end{equation}
%
We can also denote the envelope by \( A^\ast(s,r) = A(s, \W^\ast(s,r) \)
%
\begin{equation*}
\begin{aligned}[t]
  A^\ast(\lambda s, \lambda r) &= A^\ast(s,r), 
  \qquad \text{or} \qquad
  s \pdiff{A^\ast}{s} + r \pdiff{A^\ast}{r} &= 0.
\end{aligned}
\end{equation*}
%
This means that the topology of the optimal model and the envelope are fixed along any of the ``rays'' of constant \(r/s\) in the \((s,r)\)-plane, as depicted in \cref{fig:planes}(\ref{fig:scaling_plane}). 
The transition probabilities of the optimal model obey a simple scaling relation along any such ray.

@@

We saw in \cref{sec:serialarea} that the optimal models at \(s = 0\) are those with the serial topology with \(\eq\) concentrated symmetrically at the two end states.
The set of serial models with such a \(\eq\) are the limit as \(\delta \rightarrow 0\) of
%
\begin{equation}\label{eq:areamodels}
\begin{aligned}
  \W\pot_{ii+1} = \W\dep_{i+1i} = q_i, 
  \text{ for } 2 \leq i \leq M-2,
  \qquad
  \W\pot_{12} &= \delta, \quad &
  \W\pot_{M-1M} &= q_{M-1}, \\
  \W\dep_{21} &= q_1, &
  \W\dep_{MM-1} &= \delta.
\end{aligned}
\end{equation}
%
Where \(0 \leq q_i \leq rf^\nu\) for both \(\nu\).
We have an \(M-1\) parameter space of maxima.
Using results from \cref{sec:serialarea}:
%
\begin{equation*}
% \begin{aligned}
  \eqm_i = \frac{\delta}{2z} 
  \text{ for } 2 \leq i \leq M-1,
  \qquad
  \eqm_1 = \eqm_M = \frac{1}{2z},
  \text{ where } z = 1 + \frac{\delta (M-2)}{2},
  \qquad
  c_i = 2i - M - 1.
% \end{aligned}
\end{equation*}
%
We can find the other vectors of interest by:
solving \(\frg \etw = (\I - \onev \eq) \w \) for \(\etw\),
computing \(\mathbf{a} = - \enc \etw\),
and solving \(\frg \thb = (\I - \onev \eq) \mathbf{a} \) for \(\thb\).
First, some notation:
%
\begin{equation*}
  \sum_k^{i:j} f(k) 
    \equiv \sgn(j-i) \smashoperator{\sum_{k = \min(i,j)}^{\max(i,j)-1}} f(k)
    = \begin{cases}
        \sum_{k=i}^{j-1} f(k) & i < j,\\
        - \sum_{k=j}^{i-1} f(k) & i > j.
      \end{cases}
\end{equation*}
%
Then, to leading order in small \(\delta\), we find:
%
\begin{equation}\label{eq:areavecs}
\begin{aligned}
  \etwm_i - \etwm_j &= \frac{z}{\delta} \sum_k^{i:j} \frac{1}{q_k}, & \qquad
  \thbm_i - \thbm_j &= \frac{z}{\delta} \sum_k^{j:i} \frac{2k - M}{q_k}, \\
  a_i &= \frac{1}{\eqm_i} 
    \quad \text{for } 2 \leq i \leq M-1, &
  a_1 &= a_M = 1.
\end{aligned}
\end{equation}
%
Now we can use \cref{eq:kktsol} to calculate the KKT multipliers
%
\begin{equation}\label{eq:areaKKT}
\begin{aligned}
    \kktm^\nu_{ij} = \sum_k^{i:j} \frac{k - i - \delta\dep_\nu}{q_k}
      \quad \text{ for } 2 \leq i \leq M-1, \qquad
    \kktm^\nu_{1j} &= \frac{1}{\delta} \sum_{k=1}^{j-1} \frac{k - 1 - \delta\dep_\nu}{q_k},
      \\
    \kktm^\nu_{Mj} &= \frac{1}{\delta} \sum_{k=j}^{M-1} \frac{M - k + \delta\dep_\nu}{q_k}.
\end{aligned}
\end{equation}
%
They are all nonnegative, as the numerator is only negative when \(i > j\) and the negative of the sum is used.
There are only two ways a KKT multiplier can be zero. 
If the sum is empty (\(j = i\)) or it contains exactly one term that is zero (\( j = i + (-1)^\nu \)).
It is impossible for more than one term in the sum to be zero.
Those transitions ddefine the serial topology.

>>>>

This proof follows that of \cite{Lawler1988cheeger}.
We can define two inner-products that will be of use.
The first is the usual inner product of two vectors (or functions on state space in the continuum case) using the steady state distribution as the measure:
%
\begin{equation}\label{eq:innereq}
  \av{\mathbf{u}, \mathbf{v}} = \sum_i \eqm_i u_i^\ast v_i ,
  \qquad
  \nrm{\mathbf{v}} = \sqrt{\av{\mathbf{v}, \mathbf{v}}} .
\end{equation}
%
The second is an inner product of rank 2 tensors (functions of pairs of points on state space) using the steady-state flux as the measure:
%
\begin{equation}\label{eq:innerflux}
  \aav{\mathbf{F}, \mathbf{G}} = \sum_{i \neq j} \Fm_{ij} F_{ij}^\ast G_{ij} ,
  \qquad
  \nnrm{\mathbf{F}} = \sqrt{\aav{\mathbf{F}, \mathbf{F}}}.
\end{equation}
%
The omission of the diagonal terms leaves only non-negative terms.
We can think of \cref{eq:innereq} as the inner product in node-space with metric \( \diag(\eq) \), 
and \cref{eq:innerflux} as the inner product in edge-space with metric \( \diag(\F) \).
The edge-space metric is positive-definite if the sum in \cref{eq:innerflux} is restricted to the edges that have transitions.

@@

The other quantity of interest in \cref{eq:snrcoeffs} is the amplitude of the eigenmodes.
In analogy, we define the following functional (for real \(\mathbf{u}\))
%
\begin{equation*}
  \CI(\mathbf{u}) = \frac{(\eq \enc \mathbf{u}) (\mathbf{u}^\dag \w)}{r\mathbf{u}^\dag \mathbf{u}}
      = \frac{\sum_{i \neq j} \eqm_i \encm_{ij} (u_j - u_i) \sum_k \eqm_k u_k \wm_k} 
             {r\sum_{l} \eqm_l u_l^2} .
\end{equation*}
%
Following \cite{Lawler1988cheeger}, we can write
%
\begin{equation*}
  u_j - u_i = \intd[_{u_i}^{u_j}]{\alpha} 
      = \intd[_{-\infty}^{\infty}]{\alpha} 
        \brk{\vartheta(u_j - \alpha) \vartheta(\alpha - u_i) 
           - \vartheta(u_i - \alpha) \vartheta(\alpha - u_j)} ,
\end{equation*}
%
where \(\vartheta\) is the Heaviside step-function.
Then 
%
\begin{equation*}
  \begin{aligned}[t]
  \eq \enc \mathbf{u} &= \intd[_{-\infty}^{\infty}]{\alpha}  \brk{
        \Fm\pot_{\CA_\alpha \CA_\alpha^c} - \Fm\dep_{\CA_\alpha \CA_\alpha^c} 
      - \Fm\pot_{\CA_\alpha^c \CA_\alpha} + \Fm\dep_{\CA_\alpha^c \CA_\alpha} }
  %
  \; \text{where} 
    \\
    &= \intd[_{-\infty}^{\infty}]{\alpha}  \brk{
        \Fm_{\CA_\alpha \CA_\alpha^c} + \Fm_{\CA_\alpha^c \CA_\alpha} 
        - 2\Fm\dep_{\CA_\alpha \CA_\alpha^c} - 2\Fm\pot_{\CA_\alpha^c \CA_\alpha} } ,
  \end{aligned}
  %
  \begin{aligned}[t]
    \CA_\alpha &= \set{i}{u_i < \alpha} , 
    \\
    \Fm^\nu_{\CA \CB} &= \sum_{\substack{i \in \CA \\ j \in \CB}} 
                        \eqm_i \Wm^\nu_{ij} .
  \end{aligned}
\end{equation*}
%
Note that the first two terms of the integrand are of the form \( \Fm_{\CA \CB} = \Fm_{\CA \CB}\pot + \Fm_{\CA \CB}\dep \).
In exactly the same way as the proof of \cref{eq:initbnd}, we can argue that the two terms with the negative signs are indeed non-positive as they do not involve the negative diagonal elements of~\( \Fm_{ij}^\nu \).
Therefore the other terms provide an upper bound:
%
\begin{equation*}
\begin{aligned}
  \CI(\mathbf{u}) 
      &\leq \frac{\av{\mathbf{u},\w}}{r\nrm{\mathbf{u}}^2} 
        \intd[_{-\infty}^{\infty}]{\alpha}  \brk{
        \Fm_{\CA_\alpha \CA_\alpha^c} + \Fm_{\CA_\alpha^c \CA_\alpha} } 
  \\  &= \frac{\av{\mathbf{u},\w}}{r\nrm{\mathbf{u}}^2} 
        \intd[_{-\infty}^{\infty}]{\alpha} \sum_{i \neq j} \Fm_{ij} \brk{
            \vartheta(u_j - \alpha) \vartheta(\alpha - u_i) 
          + \vartheta(u_i - \alpha) \vartheta(\alpha - u_j)
        }
  \\  &= \frac{\av{\mathbf{u},\w}}{r\nrm{\mathbf{u}}^2} 
      \sum_{i \neq j} \Fm_{ij} \abs{u_j - u_i}
  \\  &= \frac{\av{\mathbf{u},\w} \aav{\abs{\boldsymbol{\Delta}\mathbf{u}}, \onem}}
              {r\nrm{\mathbf{u}}^2}.
\end{aligned}
\end{equation*}
%
Now we can apply the Cauchy-Schwartz inequality to both factors to find
%
\begin{equation*}
  \CI(\mathbf{u}) \leq \frac{\nrm{\w} \nnrm{\boldsymbol{\Delta}\mathbf{u}} \nnrm{\onem}}
                            {r\nrm{\mathbf{u}}} .
\end{equation*}
%
We also have \( \nrm{\w}^2 = \sum_i \eqm_i \wm_i^2 = 1 \).
Furthermore, using \cref{eq:constri}, \( \nnrm{\onem}^2 = \sum_{i \neq j} \eqm_i \frgm_{ij} \leq r \sum_i \eqm_i = r\).
Then, remembering \cref{eq:numrange},
%
\begin{equation}\label{eq:sqrtnumrange}
  \CI(\mathbf{u}) \leq \sqrt{\frac{2\av{\mathbf{u}, -\frg \mathbf{u}}}{r\av{\mathbf{u}, \mathbf{u}}}}
      = \sqrt{2\mathcal{W}(-\frg;\mathbf{u})/r}.
\end{equation}
%
Now, simply choosing \(\mathbf{u} = \evr_a \), one of the orthonormal eigenvectors, we find a bound on the eigenmodes:
%
\begin{equation}\label{eq:sqrtbound}
  \text{assuming} \quad 
  \eqm_i \frgm_{ij} = \eqm_j \frgm_{ji},
  \qquad \text{we have} \quad
  \CI_a \sqrt{r\tau_a} \leq \sqrt{2}.
\end{equation}
%
Numerical maximisation of the quantity above produces the same result as in \cref{sec:initial}:
the two-state model with deterministic transitions.
This model has \(\CI_1 = \tau_1 = 1\), suggesting that there's a factor of \(\sqrt{2}\) looseness in this bound.

@@

Suppose there are two states, \(i,j\), with nonzero transitions from \(m\) for one \(\nu\).
Both must statisfy \cref{eq:oneexit} for the same \(\kktm^\nu_{mm}\), so
%
\begin{equation}\label{eq:twoexits}
  \Wm^\nu_{mi}, \Wm^\nu_{mj} > 0
  \qquad \implies \qquad
  \dgnm^\nu_{m,i} = \dgnm^\nu_{m,j}
  \quad \text{or}\quad
  \dgnm_{m,i} - \dgnm_{m,j} = - (-1)^\nu (\etwm_i - \etwm_j).
\end{equation}
%
Furthermore, for another state \(k\) with zero transition rate from \(m\) for the same \(\nu\):
%
\begin{equation}\label{eq:nonexit}
\begin{aligned}
  \Wm^\nu_{mi} > 0, \Wm^\nu_{mk} = 0 
  \;\;\; \implies \;\;\;
  \kktm^\nu_{mk} = \eqm_m (\dgnm^\nu_{m,k} - \dgnm^\nu_{m,i})
  \;\; \implies \;\;
  \dgnm^\nu_{m,k} &\geq \dgnm^\nu_{m,i} \\
  \text{or}\quad
  \dgnm_{m,k} - \dgnm_{m,i} &\geq - (-1)^\nu (\etwm_k - \etwm_i).
\end{aligned}
\end{equation}
%
If both \(\nu\) have nonzero rates from \(m\) to \(i\) and \(j\):
%
\begin{equation}\label{eq:twodoubleexits}
  \Wm\pot_{mi}, \Wm\dep_{mi}, \Wm\pot_{mj}, \Wm\dep_{mj} > 0
  \qquad \implies \qquad
  \dgnm_{m,i} = \dgnm_{m,j} 
  \quad \text{and} \quad
  \etwm_i = \etwm_j.
\end{equation}
%
The consequences of these inequalities are depicted in \cref{fig:planes}(\ref{fig:triangle_spacing}).
We see that \(n\) must be a global minimum of \(\dgn_m\), and if there are multiple transitions of a single type (potentiating or depressing) the targets are constrained to lie on a single line.
Even if there is no state \(n\) that is the target of both potentiating and depressing transitions, there will be two lines at \(45^\circ \) passing through any targets of potentiating/depressing transitions that demarcate the forbidden zone.

@@

\begin{figure}[ht]
\begin{center}
\begin{myenuma}
  \item \label{fig:triangle_spacing} \aligntop{ \hspace{-2em}%
  \includegraphics[width=0.4\linewidth]{spacing_triangle.svg}}
  \hspace{0.2em}
  \item \label{fig:scaling_plane}\aligntop{ \hspace{-1em}%
  \includegraphics[width=0.4\linewidth]{scaling_plane.svg}}
\end{myenuma}
\caption[Depicting constraints on topology]{Depicting constraints on topology.
(\ref*{fig:triangle_spacing}) The consequences of \cref{eq:oneexit,eq:doubleexit,eq:twoexits,eq:twodoubleexits,eq:nonexit} in the \( (\dgnm_{m,l},-\etwm_l) \)-plane, 
  with \(i\) a state with a nonzero depressing transition from \(m\), 
  \(j\) a state with a nonzero potentiating transition, 
  \(n\) a state with both depressing and potentiating transitions 
  and \(k\) a state with no nonzero transitions from \(m\).
  The positions of \(m\) and \(k\) are not yet constrained, other than being in the upper quadrant \wrt \(n\).
(\ref*{fig:scaling_plane}) Scaling of optimal models in the \((s,r)\)-plane.
  Optimal models on grey ``rays'' are related by scaling, and therefore have the same topology.
  The line at \(r = r_0\) is where we need the solutions.
  The line at \(s = \epsilon\) is close to where we can determine the topology.
\label{fig:planes}}
\end{center}
\end{figure}

@@

We have the following invariances:
%
\begin{equation*}
  A(\lambda s, \lambda \W) = A(s, \W),
  \qquad
  \CL(\lambda s, \lambda r, \lambda \W, \lambda\inv \kkt) = \CL(s, r, \W, \kkt).
\end{equation*}
%
or in differential form:
%
\begin{equation*}
  s \pdiffc<\Wm>{A}{s} + \sum_{\nu i j} \Wm^\nu_{ij} \pdiffc<s>{A}{\Wm^\nu_{ij}} 
  = 
  s \pdiffc<r \Wm \mathrlap{\kktm}>{\CL}{s} 
    + r \pdiffc<s \Wm \mathrlap{\kktm}>{\CL}{r} 
    + \sum_{\nu i j} \brk{\Wm^\nu_{ij} \pdiffc<s r \mathrlap{\kktm}>{\CL}{\Wm^\nu_{ij}} 
                      - \kktm^\nu_{ij} \pdiffc<s r {\Wm}>{\CL}{\kktm^\nu_{ij}}} = 0.
\end{equation*}
%
In \cref{eq:kktsol} we have an expression for the KKT multipliers at the maximum in terms of \(s\) and \(\W \).
We denote this expression by \( \kkt^\ddagger(s, \W) \).
If we denote the solution to the maximisation problem by \(\W^\ast(s,r), \kkt^\ast(s,r)\), we have \( \kkt^\ast(s,r) = \kkt^\ddagger(s, \W^\ast(s,r)) \) and
%
\begin{equation}\label{eq:optmodelscale}
\begin{aligned}
  \kkt^\ddagger(\lambda s, \lambda \W) 
    &= \lambda\inv \kkt^\ddagger(s, \W), &
  \qquad \text{or} \qquad
  s \pdiffc<\Wm>{\kkt^\ddagger}{s} 
      + \sum_{\nu i j} \Wm^\nu_{ij} \pdiffc<s>{\kkt^\ddagger}{\Wm^\nu_{ij}} 
    &= -\kkt^\ddagger, \\
  \W^\ast(\lambda s, \lambda r) 
    &= \lambda \W^\ast(s,r), &
  s \pdiffc<r>{\W^\ast}{s} + r \pdiffc<s>{\W^\ast}{r} 
    &= \W^\ast, \\
  \kkt^\ast(\lambda s, \lambda r) &= \lambda\inv \kkt^\ast(s,r), &
  s \pdiffc<r>{\kkt^\ast}{s} + r \pdiffc<s>{\kkt^\ast}{r} 
    &= -\kkt^\ast.
\end{aligned}
\end{equation}
%
We can also denote the envelope by \( A^\ast(s,r) = A(s, \W^\ast(s,r)) \)
%
\begin{equation*}
\begin{aligned}[t]
  A^\ast(\lambda s, \lambda r) &= A^\ast(s,r), 
  \qquad \text{or} \qquad
  s \pdiff{A^\ast}{s} + r \pdiff{A^\ast}{r} &= 0.
\end{aligned}
\end{equation*}
%
This means that the topology of the optimal model and the envelope are fixed along any of the ``rays'' of constant \(r/s\) in the \((s,r)\)-plane, as depicted in \cref{fig:planes}(\ref{fig:scaling_plane}). 
The transition probabilities of the optimal model obey a simple scaling relation along any such ray.

@@

We saw in \cref{sec:serialarea} that the optimal models at \(s = 0\) are those with the serial topology with \(\eq\) concentrated symmetrically at the two end states.
The set of serial models with such a \(\eq\) are the limit as \(\delta \rightarrow 0\) of
%
\begin{equation}\label{eq:areamodels}
\begin{aligned}
  \W\pot_{ii+1} = \W\dep_{i+1i} = q_i, 
  % \text{ for } 2 \leq i \leq M-2,
  \qquad
  \text{except} \quad
  \W\pot_{12} &= \delta q_1, \quad &
  % \W\pot_{M-1M} &= q_{M-1}, \\
  % \W\dep_{21} &= q_1, &
  \W\dep_{MM-1} &= \delta q_{M-1}.
\end{aligned}
\end{equation}
%
Where \(0 \leq q_i \leq rf^\nu\) for both \(\nu\).
We have an \(M-1\) parameter space of maxima.
Using results from \cref{sec:serialarea}:
%
\begin{equation*}
% \begin{aligned}
  \eqm_i = \frac{\delta}{2z} 
  \text{ for } 2 \leq i \leq M-1,
  \qquad
  \eqm_1 = \eqm_M = \frac{1}{2z},
  \text{ where } z = 1 + \frac{\delta (M-2)}{2},
  \qquad
  c_i = 2i - M - 1.
% \end{aligned}
\end{equation*}
%
We can find the other vectors of interest by:
solving \(\frg \etw = (\I - \onev \eq) \w \) for \(\etw\),
computing \(\mathbf{a} = - \enc \etw\),
and solving \(\frg \thb = (\I - \onev \eq) \mathbf{a} \) for \(\thb\).
First, some notation:
%
\begin{equation*}
  \sum_k^{i:j} f(k) 
    \equiv \sgn(j-i) \smashoperator{\sum_{k = \min(i,j)}^{\max(i,j)-1}} f(k)
    = \begin{cases}
        \sum_{k=i}^{j-1} f(k) & i < j,\\
        - \sum_{k=j}^{i-1} f(k) & i > j.
      \end{cases}
\end{equation*}
%
Then, to leading order in small \(\delta\), we find:
%
\begin{equation}\label{eq:areavecs}
\begin{aligned}
  \etwm_i - \etwm_j &= \frac{z}{\delta} \sum_k^{i:j} \frac{1}{q_k}, & \qquad
  \thbm_i - \thbm_j &= \frac{z}{\delta} \sum_k^{i:j} \frac{M - 2k}{q_k}, \\
  a_i &= \frac{1}{\eqm_i} 
    \quad \text{for } 2 \leq i \leq M-1, &
  a_1 &= a_M = 1, \\
  \alpha^\nu_i &= - \frac{z ((-1)^\nu c_i + 1)}{r f^\nu \delta}
    \quad \text{for } 2 \leq i \leq M-1, &
  \alpha\pot_1 &= \frac{z(M-2)}{r f^\nu q_1}, \\
  \alpha\pot_M &= \alpha\dep_1 = 0, &
  \alpha\dep_M &= \frac{z(M-2)}{r f^\nu q_{M-1}}, \\
  \omega^\nu_i &= \frac{(-1)^\nu z}{r f^\nu \delta}
    \quad \text{for } 2 \leq i \leq M-1, &
  \omega\pot_1 &= \frac{z}{r f^\nu q_1}, \\
  \omega\pot_M &= \omega\dep_1 = 0, &
  \omega\dep_M &= - \frac{z}{r f^\nu q_{M-1}}.
\end{aligned}
\end{equation}
%
Now we can use \cref{eq:kktsol} to calculate the KKT multipliers
%
\begin{equation}\label{eq:areaKKT}
\begin{aligned}
    \kktm^\nu_{ij} = \sum_k^{i:j} \frac{k - i + \delta\dep_\nu}{q_k}
      \quad \text{ for } 2 \leq i \leq M-1, \qquad
    \kktm^\nu_{1j} &= \frac{1}{\delta} \sum_{k=1}^{j-1} \frac{k - 1 + \delta\dep_\nu}{q_k},
      \\
    \kktm^\nu_{Mj} &= \frac{1}{\delta} \sum_{k=j}^{M-1} \frac{M - k - \delta\dep_\nu}{q_k}.
\end{aligned}
\end{equation}
%
They are all nonnegative, as the numerator is only negative when \(i > j\) and the negative of the sum is used.
There are only two ways one of these KKT multipliers could be zero: 
the sum is empty; 
the sum contains exactly one term and that term is zero
(It is impossible for more than one term in the sum to be zero).
The first occurs for \(j = i\), the second for \( j = i + (-1)^\nu \).
Those transitions define the serial topology.
